A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-23 paper/fascd.tex
Some edits; not sure if I want to offer these yet

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints, namely arbitrary upper and lower obstacles, are allowed.
\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.  The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

Prior multilevel algorithms for VIs have various substantial limitations compared to the FASCD framework.  The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.  While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.  The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.  The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).  The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.  Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.  No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.  However, our implementation exceeds monotone multigrid performance, including on a problem used to demonstrate the benefits of that method (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).

>>>>

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.  The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

Prior multilevel algorithms for VIs have various substantial limitations compared to the FASCD framework.  The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.  While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.  The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.  The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).  The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.  Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.  \pef{My current understanding is that this refers to \emph{truncated} monotone multigrid; monotone multigrid just means using monotone injection operators appropriately. Let me double-check. And my understanding is that the `modification' of the basis functions is just zeroing the basis functions for vertices in the active set, so you don't have to modify the basis functions, just zero some rows/columns of the matrix.} No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.  However, our implementation exceeds monotone multigrid performance, including on a problem used to demonstrate the benefits of that method (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
